---
title: "STATISTICAL COMPUTING PROJECT"
subtitle: Adem VAROL - 200709078 / Wael KABOUK - 190709723
#author: "Adem VAROL - 200709078 / Wael KABOUK - 190709723"
output:
  html_document: default
  pdf_document: default
font-family: Gill Sans
date: "`r format(Sys.time(), '%d %B %Y')`"
---

```{r setup, include=FALSE }
knitr::opts_chunk$set(echo = TRUE)
```

## 1. (3p) Data:
Please find your original data set or data sets; and describe your data in the first step.

The real data of OAL and YEAL schools were used in order to examine the success of the two schools according to the GPA of the students and the status of their families (for example, parents divorced, mother died, father died, etc.).

GitHub Link: https://github.com/200709078/SC_PROJECT

## 2. (3p) Exploratory and descriptive data analysis:
Use “Exploratory and descriptive data analysis”. Talk about your categorical and quantitative data or your ordinal variables etc. Write down your comments.
```{r}
## install.packages("readxl")
library("readxl")
dFile <- read_excel("SC_DATA.xls")
names(dFile)
```
ORDINAL VARIABLE:   
NUMBER (1470, 1471, ...)    
FULLNAME (Adem VAROL, ... , Wael KABOUK)    
  
CATEGORICAL VARIABLES:    
SCHOOL (OAL, YEAL)    
CLASS (9A, 9B, 9C, ...)   
GENDER (1 : Male, 0 : Female)   
STAY (1 : Stay in the town, 0 : Stay in the village)    
MAF_PS (Mother and father are primary school graduates, 1 : Yes, 0 : No)    
CHILD_1 (1 : One Kid, 0 : Two Kids)   
CHILDREN_3 (1 : Three Kids, 0 : More than three children)   
MAF_DIV (Mother and father divided, 1 : Yes, 0 : No)    
MOTHER_DIED (1 : Yes, 0 : No)   
FATHER_DIED (1 : Yes, 0 : No)   
FAMILY_PATIENT (Has a family member who is sick, 1 : Yes, 0 : No)   
STUDENT_PATIENT (The student has a sick,  1 : Yes, 0 : No)    
  
QUANTITATIVE VARIABLE:    
GPA (Grade point average)   
```{r}
summary(dFile$GPA)
## head(dFile)
```
## 3. (3p) Data Visualization:
Use at least 4 different “data visualization techniques” and talk about your data further (distribution, outliers, variability, etc). Use at least 2 of the visualizations to compare two groups (like female/male; smoker/non-smoker etc).

```{r}
library(MASS)
library(ggplot2)

barplot(dFile$GPA, main = 'BAR PLOT FOR GPA', xlab = 'STUDENTS', ylab = "GRADES", horiz = FALSE, ylim = c(0,100))

hist(dFile$GPA, main ="HISTOGRAM PLOT FOR GPA", xlab ="GPA", xlim = c(0, 100), col ="blue", freq = TRUE, breaks = 10)

plot(density(subset(dFile, GENDER=="0")$GPA), xlab = "GPA OF GIRLS", main = "HISTOGRAM PLOT FOR GIRLS",col="pink")
plot(density(subset(dFile, GENDER=="1")$GPA), xlab = "GPA OF BOYS", main = "HISTOGRAM PLOT FOR BOYS",col="blue")

ggplot(dFile, aes(x=GPA)) + geom_density(col="red")

qqnorm(dFile$GPA, pch = 1, frame = TRUE, main ="NORMAL Q-Q PLOT FOR GPA")
# qqline(dFile$GPA, col = "steelblue", lwd = 2)

dataOAL <- subset(dFile, SCHOOL=="OAL")
dataYEAL <- subset(dFile, SCHOOL=="YEAL")
boxplot(main="BOX PLOT FOR GPA", dataOAL$GPA, dataYEAL$GPA, ylab ="GPA", col="steelblue")

TABL=table(dFile$GENDER)
barplot(TABL,beside = T, legend = T)

```

## 4. (3p) Central Limit Theorem:
Choose one of your variables. Write down your own function which proves the central limit theorem (CLT): show that whatever the distribution is, the sampling data will be distributed normally. (have a look at the example in Statistics_Lecture_5, page9)
```{r}
library(shiny)

hist(dFile$GPA, main ="Histogram Plot for GPA", xlab ="GPA", xlim = c(0, 100),
    col ="blue", freq = TRUE, breaks = 10)

shapiro.test(dFile$GPA)

#H0: Data is Normally Distributed
#H1: Data is not Normally Distributed    alpha=0.05

#If p ≤ 0.05: then the null hypothesis can be rejected 
#If p > 0.05: then the null hypothesis cannot be rejected 
# So p is less than 0.05 our data is not normally distributed.

```
## 5. (3p) Confidence Intervals:
Build 2 confidence intervals and make “clear comments” about your findings.

```{r}
#t-test
gpa.mean <- mean(dFile$GPA)
gpa.sd <- sd(dFile$GPA)
gpa.se <- gpa.sd/sqrt(length(dFile))

t.test(dFile$GPA,mu=2000)


#Confidence interval 
t.score1 <-  qt(p=0.05, df=80,lower.tail=F)

error <- t.score1 * gpa.se
lower.bound <- gpa.mean  - error
upper.bound <- gpa.mean  + error
sprintf('%f < x < %f',lower.bound, upper.bound)
# print(c(lower.bound, upper.bound))

dataA <- subset(dFile, lower.bound < GPA & GPA<upper.bound)

#1 — Hypothesis for first Fisher’s Exact Test (Alternative Two Sided):
#Ho: The odds ratio is equal to 1
#Ha: The odds ratio is not equal to 1
#2 — Hypothesis for first Fisher’s Exact Test (Alternative Less):
#Ho: The odds ratio is equal to or greater than 1
#Ha: The odds ratio is less than 1
#3 — Hypothesis for first Fisher’s Exact Test (Alternative Greater):
#Ho: The odds ratio is equal to or less than 1
#Ha: The odds ratio is greater than 1
```

## 6. (3p) Transformation:
Try to make one transformation (log transformation, Box-Cok transformation, etc) for one of your quantitative variables, which is not normally distributed; but will be normal or more normal, after the transformation.
```{r}

withoutoutliers <- subset(dFile, 10<GPA & GPA<90)

#log transform
dFile.log <- transform(withoutoutliers, GPA.log=log(GPA))

#histograms
hist(dFile.log$GPA.log, col="blue",xlab = "GPA" ,main = "Histogram Plot for GPA",breaks = 15)

#density plots with basic R
d.GPA <- density(dFile.log$GPA.log)
plot(d.GPA, xlab = "GPA",main = "Histogram Plot for GPA",col="blue")

#density plots with ggplot2
ggplot(dFile.log, aes(x=GPA.log)) + geom_density()

#qqplots
qqnorm(dFile.log$GPA.log, pch = 1, frame = FALSE)
#qqline(dFile.log$GPA.log, col = "steelblue", lwd = 2,xlab = "GPA")

#boxplots
boxplot(dFile.log$GPA.log)
```

## 7. (2p every item) Single t-test (Welch t-test or Wilcoxon rank-sum test)
Implement a single t-test for one of your “normally or not-normally distributed” variable:

a. Aim
In words, what is your objective here?

**Checking whether the average GPA of the whole students is equal or greater than 70 **

b. Hypothesis and level of significance:
Write your hypothesis in scientific form and determine the level of significance.

**H0: mu=70, H1: mu < 70, alpha = 0.05**

c. Assumption Check:
Check the required assumptions and “comment on each of them is a must!”.

**Assuming the data is not normally distributed, equal variance **

d. Indicate “which test you choose” “for what reason”

**Single t-test,assumptions above **

e. Result:
Give the output of the test and write down the result (ex: since p value is less than alpha, I reject the null hypothesis).

**Since p-value is less than alpha we reject the null hypothesis**

```{r}
boxplot(dFile$GPA)

TEST<- t.test(dFile$GPA, mu=70, alt = "less", conf = 0.95)
TEST$null.value
TEST$alternative
TEST$conf.int
TEST$p.value

```

f. Conclusion:
You got your result in item e. Write down the conclusion of your result, in such a way that, the reader who doesn’t know any statistics can understand your findings.

**We are 95% confident that the average GPA of the students is equal or greater than 70**

g. What can be Type-1 and Type-2 error here?

**Type-1 Error: when rejecting the null hypothesis while it is in fact true, Type-2 Error: When not rejecting the null hypothesis while it is in fact false**

## 8. (2p every item) Paired t-test:

```{r}
gpaOAL <- dFile$GPA[1:451]
gpaYEAL <- dFile$GPA[452:902]

boxplot(gpaOAL,gpaYEAL)

plot(gpaOAL, gpaYEAL, abline(a=0, b=1))

#H0: Mean difference in SBP is 0
# two-sided test

t.test(gpaOAL, gpaYEAL, mu=0, alt="two.sided", paired = T, conf = 0.99)

```

a. Aim
In words, what is your objective here?

**To compare the difference between two selected groups in the same population**

b. Hypothesis and level of significance:
Write your hypothesis in scientific form and determine the level of significance.

**H0: mu = 0, H1: mu != 0, alpha = 0.01**

c. Assumption Check:
Tell why you use the paired t test, and Check the required assumptions. “comment on each of them is a must!”.

**The variables are metric and the difference of the paired values is normally distributed**

d. Result:
Give the output of the test and write down the result (ex: since p value is less than alpha, I reject the null hypothesis).

**Since p value is less than alpha, I reject the null hypothesis**

e. Conclusion:
You got your result in item d. Write down the conclusion of your result, in such a way that, the reader who doesn’t know any statistics can understand your findings.

**We are 99% confident that there is no significant difference between the means of the two selected groups of the students GPAs**


## 9. (2p every item) Fisher’s exact test for count data

```{r}
TAB = table(dFile$GENDER,dFile$MAF_PS)

barplot(TAB,beside = T, legend = T)

TESTFISHER = fisher.test(TAB, conf.int = T, conf.level = 0.99)
TESTFISHER

```

a. Aim    
In words, what is your objective here?    

**We applied this test to compare gender and parents whose parents were primary school graduates.**   

b. Hypothesis and level of significance:
Write your hypothesis in scientific form and determine the level of significance.   

**Hypothesis: There is a difference between gender groups (female=0/male=1) in terms of parents being primary school graduates.**   
**#H0: The odds ratio is equal to or greater than 1**   
**#H1: The odds ratio is less than 1**   
  
c. Result:    
Give the output of the test and write down the result (since p value is less than alpha, I reject the null hypothesis).   

**since p value (p-value = 0.2324) is less than 1, I reject the null hypothesis.**    
  
d. Conclusion:    
You got your result in item c. Write down the conclusion of your result, in such a way that, the reader who doesn’t know any statistics can understand your findings.    

**We can say that the status of parents being primary school graduates has the same effect on gender groups.**    
  
e. Odds Ratio:    
Comment about the odds ratio, what does it indicate?    

**Odds ratio shows us the relationship between the probability of a situation occurring and the probability of it not happening. The calculation is quite simple, the probability of occurrence of the curious situation is divided by the probability of not happening. Odds ratio (0.8566012) tells us that; The effect of gender on parents being primary school graduates is 0.86 times.**   
  
## 10. (2p every item) ANOVA and Tukey Test   

```{r}
# ANOVA TEST
group1 <- dFile$GPA[1:200]   # OAL first 200 students
group2 <- dFile$GPA[201:400] # OAL second 200 students
group3 <- dFile$GPA[452:651] # YEAL first 200 students
group4 <- dFile$GPA[652:851] # YEAL second 200 students

combined_groups <- data.frame(cbind(group1,group2,group3,group4))
boxplot(combined_groups)
summary(combined_groups)

stacked_groups <- stack(combined_groups)
#stacked_groups
boxplot(stacked_groups$values)
summary(stacked_groups)

anova_result <- aov(values~ind, data=stacked_groups)
summary(anova_result)

# F(3, 796) = 159.9, p=2e-16 < 0.05

# TUKEY TEST
# H0: All means of genders are equal
# H0: There is no difference in gender
# H0: There is no interaction between gender

TukeyHSD (anova_result)
model.tables(anova_result,"mean")

barplot(as.matrix(combined_groups), beside = TRUE, legend.text = TRUE, ylim = c(0,100), xlab = "GROUPS", ylab = "GPA", main = "BARPLOT FOR ANOVA")

```

a. Aim    
In words, what is your objective here?    
  
**To test whether there is a statistically significant difference according to the means of 4 different selected groups.**    
  
b. Hypothesis and level of significance:    
Write your hypothesis in scientific form and determine the level of significance.   
  
**H0: There is no statistically significant difference according to the means between the groups.**   
**H1: There is a statistically significant difference according to the means between the groups.**    
  
c. Assumption Check:    
Check the required assumptions. “comment on each of them is a must!”.   

**There is a statistically significant difference between group4-group3 and group2-group1 as the p_adj value is different from 0. There is no significant difference between the others.**    
  
d. Result of ANOVA:   
Give the output of the test and write down the result (ex:since p value is less than alpha, I reject the null hypothesis)   
  
**Since p=2e-16 is less than 0.05, I reject the null hypothesis.**    
  
e. Conclusion of ANOVA:   
You got your result in item d. Write down the conclusion of your result, in such a way that, the reader who doesn’t know any statistics can understand your findings.    
  
**Since there is a great difference between the averages of group4-group3 and group2-group1, these groups are not statistically similar. So this group does not have the same level of success.**    
  
f. Result of Tukey:   
Give the output of the test and write down the result (ex:since p value is less than alpha, I reject the null hypothesis)   
  
**Since p=2e-16 is less than 0.05, I reject the null hypothesis for Tukey test.**   
  
g. Conclusion of Tukey:   
You got your result in item f. Write down the conclusion of your result, in such a way that, the reader who doesn’t know any statistics can understand your findings.    
  
**According to the Tukey test, since there is a large difference between the averages of group4-group3 and group2-group1, these groups are not statistically similar. In other words, these groups do not have the same level of success.**    
  
## 11. (2p every item) Multiple Linear Regression   
  
```{r}
modelMLR <- lm(dFile$GPA ~ dFile$GENDER + dFile$STAY)
summary(modelMLR)
# Approximately (Multiple R-squared:  0.01244) 1% of variation in GPA can be explained by our modelMLR (GENDER + STAY)

# This tests the Null Hypothesis that all the modelMLR coefficients are 0 (F-statistic: 7.027 on 2 and 1116 DF,  p-value: 0.0009272)

# H0: Bgender=Bstay=0

# Gives on idea of how far observed GPA (y values) are from the predicted or fitted GPA (Residual standard error: 19.61 on 1116 degrees of freedom)

# Calculate Pearson's correlation between Gender and Stay
cor(dFile$GENDER, dFile$STAY, method = "pearson")

# Ask for confidence intervals for the model coefficients
confint(modelMLR, conf.level=0.95)

plot(modelMLR)

```
  
a. Aim    
In words, what is your objective here?    
  
**We examined the linear correlation between GPA dependent variable and GENDER and STAY independent variables with Multiple Linear Regression.**    
  
b. Regression Equation:   
Multiple linear regression (MLR) is a statistical technique that uses several explanatory variables to predict the outcome of a response variable. Which ones are your explanatory variables and which one is your response variable? Write down the equation of your regression using those variables.
  
**That is, independent variables are X, dependent variables are Y, and Y=XB+E.**    
**# F(59.2752, 60.790), p=2e-16 < 0.05**    
  
c. Hypothesis and level of significance:    
Write your hypothesis in scientific form and determine the level of significance.   
  
**H0: GENDER and STAY independent variables are linearly dependent with GPA dependent variable.**   
**H1: There is no Multiple Linear Regression between GENDER and STAY independent variables and GPA dependent variable.**    
  
d. Find the Best Model:   
Use step function and find the best model, describe the reason which makes it the best one.   
  
**Since p-value= 0.0009272 <0.5, the best model is the step wise model.**   
  
e. Assumption Check:    
Check the required assumptions, “comment on each of them is a must!”.   
  
**(F-statistic: 7.027 on 2 and 1116 DF, p-value: 0.0009272) Since p-value<0.5, there is a linear dependence between the variables.**    
  
f. Result:    
Give the output of the best model and write down the result.    
  
**Since p-value= 0.0009272 <0.5, the best model is the step wise model.**   
  
g. Conclusion:    
You got your result in item f. Write down the conclusion of your result, in such a way that, the reader who doesn’t know any statistics can understand your findings.    
  
**Gender and the place where the student lives do not affect the GPA.**   
  
h. Prediction:    
Think of a new X variable, or variables (just make it up). Use your best model in order to predict the new Y variable for those X variables.    
  
**If a new variable such as the student's weight was added to our dataset, this would not have an effect on the GPA. We could use the stepwise model for this.**
  
## THANKS FOR ALL ##








